Boosted Mirror Descent, Frank-Wolfe, and Herding

### 2013-10-04 icml/introduction.tex
first draft of introduction

\documentclass[paper.tex]{subfiles}
\begin{document}
\section{Boosted Mirror Descent}
\label{sec:algorithm}

>>>>

\documentclass[paper.tex]{subfiles}
\begin{document}

\section{Introduction} 
\label{sec:intro}

\NA{JHH: Need to add in all these citations}

The herding algorithm, which was first presented in \citet{Welling:2009a}, has received increasing attention since its introduction. Herding was originally introduced as a deterministic method for performing joint learning and inference in Markov random field (MRFs). This derivation was based on performing fixed step size gradient descent on the zero temperature limit the maximum entropy optimization problem for an MRF. Herding for MRFs was further explored in \citet{Welling:2009a,Gelfand:2010,Bornn:2013}.

Let $\sX$ be an observation space, $\phi : \sX \to \sB_{\Theta}$ a feature map to an inner produce space, and $\bar\phi = \bE_{x \sim p}[\phi(x)]$ the empirical moments of some (possible empirical) distribution. Then the herding algorithm is given by the follow simple procedure:
\[
x_{t} &\in \argmin_{x \in \sX} \ip{\theta_{t}}{\phi(x)} \\
\theta_{t} &= \theta_{t-1}  + \bar \phi - \phi(x_{t}).
\]

An alternative interpretation of herding \citep{Chen:2010a,Huszar:2012,Bach:2012a} is as a moment matching algorithm. In particular, it generates a set of pseudo-samples $\{ x_{t} \}$  which match the moments $\bar\phi$. That is, the error $\sE_{T}^{2} \eqdef \| \bar\phi - T^{-1}\sum_{t=1}^{T} \phi(x_{t})\|^{2} \to 0$ as $T \to \infty$.  If $\sB_{\Theta}$ is a reproducing kernel Hilbert space generated by $\phi$, $\sE_{T}^{2}$ is equivalent to the maximum mean discrepancy (MMD) between $p$ and the empirical distribution $\hat p_{T} \eqdef T^{-1}\sum_{t=1}^{T} \delta_{x_{t}}$ \citep{Huszar:2012}:
\(
\sE_{T}^{2} &= \text{MMD}^{2}(p, \hat p_{T})  \\
&\eqdef \sup_{\theta \in \Theta} \left|\bE_{p}\ip{\theta}{\phi(x)} - \bE_{\hat p_{T}}\ip{\theta}{\phi(x)} \right|,
\)
where $\Theta = \{ \theta \in \sB_{\Theta} \ | \  \|\theta\| = 1 \}$.

If we are interested in generating samples from $p$ (or some distribution with the same moments as $p$), herding provides an attractive alternative to standard i.i.d. sampling for two reasons. First, it is deterministic, so we do not have to worry about getting ``unlucky'' and generating a set of samples that poorly match $p$. Second, in the case where $\sB_{\theta}$ is finite dimensional, the error $\sE_{T}$ of the herding estimator converges to zero at a rate of $O(1/T^{2})$ whereas the convergence rate of i.i.d.~sampling is $O(1/T)$ \citep{Chen:2010a}. The infinite dimensional case, however, is not as well understood. \citet{Bach:2012a} recently showed that the $O(1/T^{2})$ convergence proof of \citet{Chen:2010a} does not apply in the infinite dimensional case. In addition, \citet{Bach:2012a} introduced an interpretation of herding as a special case of conditional gradient descent (\algname{CGD}) minimizing the square norm. They showed that \algname{CGD} with line search produces faster convergence rates: $O(1/T)$ in the infinite dimensional case and exponential in the finite dimensional valse. However, they found that empirically that line search had poor approximation properties since led to distributions far from the maximum entropy distribution.

In this paper, we make the following contributions:
\begin{itemize}
\item In Section \ref{sec:algorithm} we introduce a novel pair of convex algorithms: primal and dual boosted mirror descent (\algname{PBMD} and \algname{DBMD}). We show how they relate to \algname{CGD} and mirror descent (\algname{MD}). 
\item We then use \algname{DBMD} in Section \ref{sec:herding} to define a class of generalized herding algorithms. Connections with the learning interpretation are made by showing how to obtain herding as the limit of an entropically normalized case of \algname{BMD}. 
\item Section \ref{sec:proofs} presents convergence rates for $\algname{PBMD}$ and $\algname{DBMD}$. As special cases, we recover a $O(\log T/ T)$ convergence rate for herding and a $O(1/T)$ rate for a herding-like algorithm that uses unequally weighted samples. Both of these results apply to the infinite dimensional case. 
\item In Section \ref{sec:lower-bounds} we prove matching lower bounds  showing that ``sparse'' herding-like methods cannot converge faster than $O(1/T)$ in the the infinite dimensional case and that herding converges at the slower $O(\log T / T)$ rate.
\item Finally, in section \ref{sec:chen} we generalize the $O(1/T^{2})$ proof of \citet{Chen:2010a} to a class of herding-like algorithms derived from \algname{BMD}. 
\end{itemize}

\section{Boosted Mirror Descent}
\label{sec:algorithm}
